Tail Flexibility in the Degrees of Preferential Attachment Networks

### 2025-04-30 paper.qmd
edit qmd w/ added ref (in bib); rm (untrack) paper.pdf

If the power law is an adequate model to use for real degree distributions has become a heated debate, with alternatives such as [@Broido_2019] seeming to outperform the power law in some cases. Additionally, [ref] find that for many real degree distributions the body may follow a power law, the right tail does not or at least seems to be lighter than is implied by the body whilst still being heavy. One method for modelling the tail of the degree distributions of real networks is to use a discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (GPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ is defined by the survival:

@@

@@

These models that seem to outperform the power law do not however come with the added benefit of suggesting a generating mechanism for the networks in the same way the power law does. Efforts to make the preferential attachment model more flexible have not yet remedied this issue with [@krapivsky01] finding that using a sub/super-linear preference function yields a Weibull tail and a degenerate network, where one vertex eventually gains all new edges, respectively.

The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in [@rudas07] and is defined as follows:

@@

```{r fig-rec2, out.width="80%"}
#| echo: false
#| fig.cap: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
knitr::include_graphics("mcmc_plot.jpg")
```

@@

Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources. [describe networks being used]. Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in [clement] and is similar to others that have been used to model degree distributions [others].

@@

```{r fig-rec3, out.width="80%"}
#| fig-cap: Posterior estimates of paramters for real data.
#| echo: false
knitr::include_graphics("pars_plot.png")
```

>>>>

If the power law is an adequate model to use for real degree distributions has become a heated debate, with alternatives such as @Broido_2019 seeming to outperform the power law in some cases. Additionally, @lef24 find that for many real degree distributions the body may follow a power law, the right tail does not or at least seems to be lighter than is implied by the body whilst still being heavy. One method for modelling the tail of the degree distributions of real networks is to use a discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (GPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ is defined by the survival:

@@

[CL: Is the precise form of the IGPD needed here? The big picture is probably more important: extreme value methods have not been used in analysing degree distributions until recently, such as @vvvk19 and @lef24. Not using these methods essentially downweights the deviation of the vertices with large degrees, which have far more influence on the network than those with small degrees, from the power law behaviour.]

@@

@shimura12 then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$. Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,n) = 0$ then the distribution is light tailed.

[CL: The results of @shimura12 could be moved closer to Proposition 2.1. In the introduction, if you want to include @shimura12, you can mention the paper in passing, when outlining your approach, e.g. "Using the results by @shimura12, we derive the tail index / heaviness of the limiting degree distribution of the GPA model [@rudas07]. This informs us the precise class of preference functions that result in flexible heavy tail. "]

These models that seem to outperform the power law do not however come with the added benefit of suggesting a generating mechanism for the networks in the same way the power law does. Efforts to make the preferential attachment model more flexible have not yet remedied this issue with @krapivsky01 finding that using a sub/super-linear preference function yields a Weibull tail and a degenerate network, where one vertex eventually gains all new edges, respectively.

The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in @rudas07 and is defined as follows:

@@

```{r fig-rec2, out.width="80%"}
#| echo: false
#| fig.cap: Posterior estimates of parameters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
knitr::include_graphics("mcmc_plot.jpg")
```

@@

Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources. [describe networks being used]. Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in @lef24 and is similar to others that have been used to model degree distributions [others].

@@

```{r fig-rec3, out.width="80%"}
#| fig-cap: Posterior estimates of parameters for real data.
#| echo: false
knitr::include_graphics("pars_plot.png")
```

### 2025-04-30 paper.tex
edit qmd w/ added ref (in bib); rm (untrack) paper.pdf

If the power law is an adequate model to use for real degree
distributions has become a heated debate, with alternatives such as
\citep{Broido_2019} seeming to outperform the power law in some cases.
Additionally, {[}ref{]} find that for many real degree distributions the
body may follow a power law, the right tail does not or at least seems
to be lighter than is implied by the body whilst still being heavy. One
method for modelling the tail of the degree distributions of real
networks is to use a discretised variation of the generalised Pareto
distribution (GPD) usually dubbed the Integer GPD (GPD) where
\(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by the survival:

@@

@@

These models that seem to outperform the power law do not however come
with the added benefit of suggesting a generating mechanism for the
networks in the same way the power law does. Efforts to make the
preferential attachment model more flexible have not yet remedied this
issue with \citep{krapivsky01} finding that using a sub/super-linear
preference function yields a Weibull tail and a degenerate network,
where one vertex eventually gains all new edges, respectively.

The network generative model that we will be focussing on in this paper
is dubbed General Preferential Attachment in \citep{rudas07} and is
defined as follows:

@@

\caption{\label{fig-rec2}Posterior estimates of paramters for data
simulated from the proposed model with various combinations of
(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}

@@

Turning now to real data, the goal is to fit the model to the degree
distributions of real networks from various sources. {[}describe
networks being used{]}. Alongside fitting this model to the degree
distributions, we then compare the fit to that of an mixture
distribution that was used in {[}clement{]} and is similar to others
that have been used to model degree distributions {[}others{]}.

@@

\caption{\label{fig-rec3}Posterior estimates of paramters for real
data.}

>>>>

If the power law is an adequate model to use for real degree
distributions has become a heated debate, with alternatives such as
\citet{Broido_2019} seeming to outperform the power law in some cases.
Additionally, \citet{lef24} find that for many real degree distributions
the body may follow a power law, the right tail does not or at least
seems to be lighter than is implied by the body whilst still being
heavy. One method for modelling the tail of the degree distributions of
real networks is to use a discretised variation of the generalised
Pareto distribution (GPD) usually dubbed the Integer GPD (GPD) where
\(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by the survival:

@@

\citep[CL: Is the precise form of the IGPD needed here? The big picture
is probably more important: extreme value methods have not been used in
analysing degree distributions until recently, such as][ and
\citet{lef24}. Not using these methods essentially downweights the
deviation of the vertices with large degrees, which have far more
influence on the network than those with small degrees, from the power
law behaviour.]{vvvk19}

@@

\citet{shimura12} then states that if
\(\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha\) (\(\alpha>0\)),
then \(F\) is heavy tailed with \(\bar F(n) \sim n^{-\alpha}\).
Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,n) = 0\) then the
distribution is light tailed.

\citep[CL: The results of][ could be moved closer to Proposition 2.1. In
the introduction, if you want to include \citet{shimura12}, you can
mention the paper in passing, when outlining your approach, e.g.~``Using
the results by \citet{shimura12}, we derive the tail index / heaviness
of the limiting degree distribution of the GPA model \citep{rudas07}.
This informs us the precise class of preference functions that result in
flexible heavy tail.'']{shimura12}

These models that seem to outperform the power law do not however come
with the added benefit of suggesting a generating mechanism for the
networks in the same way the power law does. Efforts to make the
preferential attachment model more flexible have not yet remedied this
issue with \citet{krapivsky01} finding that using a sub/super-linear
preference function yields a Weibull tail and a degenerate network,
where one vertex eventually gains all new edges, respectively.

The network generative model that we will be focussing on in this paper
is dubbed General Preferential Attachment in \citet{rudas07} and is
defined as follows:

@@

\caption{\label{fig-rec2}Posterior estimates of parameters for data
simulated from the proposed model with various combinations of
(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}

@@

Turning now to real data, the goal is to fit the model to the degree
distributions of real networks from various sources. {[}describe
networks being used{]}. Alongside fitting this model to the degree
distributions, we then compare the fit to that of an mixture
distribution that was used in \citet{lef24} and is similar to others
that have been used to model degree distributions {[}others{]}.

@@

\caption{\label{fig-rec3}Posterior estimates of parameters for real
data.}
